\documentclass{article}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}

% Header
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{MAD 103 - Data Fundamentals - Fall 2024}
\fancyhead[R]{Instructor: Sodiq Shofoluwe  \thepage}

% Formatting for title sections
\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}

\title{Course Syllabus\\\large MAD 103: Data Fundamentals - Fall 2024}
\author{Instructor: Sodiq Shofoluwe}
\date{September 5th, 2024}

\begin{document}

\maketitle

\section*{Course Outline: MAD 103}
\textbf{Instructor:} Sodiq Shofoluwe \\
\textbf{Term:} Fall 2024

\section*{Important Dates}
\begin{itemize}
    \item September 3rd, 2024 - Semester starts
    \item September 16th, 2024 - Last day to add/withdraw for the fall semester
    \item October 14th, 2024 - Thanksgiving Holiday (College Closed)
    \item November 12th, 2024 - 2/3rd drop date (Last date to drop a class)
    \item December 9th - 13th, 2024 - Final Test Week
\end{itemize}

\section*{Course Syllabus}
\textbf{Course:} Data Fundamentals

\textit{NOTE: This outline is a guide only. Changes from week to week may occur. You will be notified via your college email address of any changes.}

\subsection*{Week-by-Week Breakdown}
\begin{description}[align=left,labelwidth=4cm,style=nextline]
    \item[Week 01] Course Introduction – XML Documents \\ Assignment \#1 \\ Quiz \#1
    \item[Week 02] Working with Data – Introduction to JSON \\ Assignment \#2 \\ Quiz \#2
    \item[Week 03] Types of Information \\ Assignment \#3 \\ Quiz \#3
    \item[Week 04] TEST \#1 (Covers week 1-3)
    \item[Week 05] Data Definition Language – CREATE, ALTER, DROP \\ Assignment \#4 \\ Quiz \#4
    \item[Week 06] Data Manipulation Language – INSERT, UPDATE, DELETE \\ Assignment \#5 \\ Quiz \#5
    \item[Week 07] Data Manipulation Language – SELECT \\ Assignment \#6 \\ Quiz \#6
    \item[Week 08] Data Manipulation Language \& Review – Working with Functions \\ Assignment \#7 \\ Quiz \#7
    \item[Week 09] Data Manipulation Language – Working with Aggregate Functions \\ Assignment \#8 \\ Quiz \#8
    \item[Week 10] TEST \#2 (Covers week 4-8)
    \item[Week 11] Data Manipulation Language – Working with Subsets of Information \\ Assignment \#9 \\ Quiz \#9
    \item[Week 12] Data Manipulation Language – Working with Multiple Tables \\ Assignment \#10 \\ Quiz \#10
    \item[Week 13] Normalization \\ Assignment \#11
    \item[Week 14] Entity Relationship Diagrams \\ Assignment \#12
    \item[Week 15] TEST \#3 (Covers week 9-14)
\end{description}

\section*{Important Policies}
\begin{itemize}
    \item \textbf{Assignments and Quizzes:} All assignments and quizzes are due by the posted due dates and times. Any assignments submitted past the posted due date and time will receive a mark of 0.
    \item \textbf{Individual Work:} These are individual assignments – no group work is allowed. All assignments must be completed by you alone. Any assignments that match will receive a mark of 0 and will be handled as per the Code of Student Rights and Responsibilities.
    \item \textbf{Generative AI:} The use of generative artificial intelligence tools and apps is strictly prohibited in all course assessments. Use of generative AI in this course is considered Academic Misconduct.
\end{itemize}
\newpage
\section*{Assessment Totals}
\begin{itemize}
    \item Assignments: 30\%
    \item Test 1: 20\%
    \item Test 2: 20\%
    \item Test 3: 20\%
    \item Quizzes: 10\%
\end{itemize}

\end{document}