\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{setspace}

% Header
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{MAD 103 - Data Fundamentals}
\fancyhead[R]{Instructor: Sodiq Shofoluwe \thepage}

\title{\textbf{TEST \#3 - Important Information}}
\author{}
\date{December 9th, 2024}

\begin{document}

\maketitle

\section*{Test Breakdown}
\begin{itemize}
    \item 2 short answer questions - 5 marks each
    \item 2 short answer questions - 3 marks each
    \item 6 short answer questions - 2 marks each
    \item Total: 10 questions - 28 marks
\end{itemize}

\textbf{Be aware of the time and the questions involved. There is a 60-minute time limit on the test.}

\section*{Important Information About Your Answers}
\begin{enumerate}
    \item When using functions, you \textbf{MUST} provide an alias to give a cleaner name.
    \item You \textbf{MUST ONLY} provide the information that is requested. For example, if a question asks for a name and phone number and you use \texttt{SELECT *}, you will lose marks.
    \item The short answer questions will ask you to create appropriate queries to meet the requirements of each question.
    \item Use the provided table to answer all questions (see link below).
    \item To answer the short answer questions, you can use a text editor to build your queries and test them using \texttt{sql.scweb.ca}.
\end{enumerate}

\section*{Do Your Own Work for the Test!}
\begin{itemize}
    \item You are restricted to using the following items for the test:
    \begin{itemize}
        \item A single 8 1/2 x 11 handwritten sheet of paper.
        \item Blackboard / Respondus LockDown Browser.
    \end{itemize}
    \item You are \textbf{NOT} allowed to have any other assistance when taking this test!
    \item Any query that will not run will receive a mark of 0.
    \item Any answers not manually built will receive a mark of 0.
    \item You are only allowed to apply the skills and knowledge that we have learned in class.
    \item Any answers that do not follow the practices and knowledge that we have learned in class will receive a mark of 0.
    \item All the work \textbf{MUST} be completed by you and only you.
    \item Any work not completed by you will be handled as per the Student Code of Rights and Responsibilities and will result in a 0 grade for this test.
\end{itemize}

\end{document}