\documentclass{article}
\usepackage{fancyhdr}

% Header
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{MAD 103 - Data Fundamentals}
\fancyhead[R]{Instructor: Sodiq Shofoluwe \thepage}

\begin{document}

\title{Test \#1 - Important Information}
\author{}
\date{September 26, 2024}
\maketitle

\section*{Instructions}
Please read the following instructions carefully:

\begin{itemize}
    \item \textbf{Test Availability}: The test will be accessible in this folder starting at 2:10 pm on September 26, 2024.
    \item \textbf{Platform}: The test will be conducted on Blackboard using the Respondus LockDown browser.
    \item \textbf{Time Limit}: 45 minutes, ending at 3:00 pm.
\end{itemize}

\section*{Test Breakdown}
\begin{itemize}
    \item \textbf{Total Questions}: 22
    \item \textbf{Question Types and Marks}:
    \begin{itemize}
        \item 20 multiple-choice, true/false questions worth 1 mark each
        \item 2 short answer questions worth 10 marks each
    \end{itemize}
    \item The short answer questions will require you to create well-formed and valid XML and JSON files.
\end{itemize}

\section*{Guidelines}
\begin{itemize}
    \item Use a text editor, browser, and JSON validator (e.g., JSONLint) to complete and validate your answers.
    \item \textbf{Remove formatting when pasting answers}.
\end{itemize}

\section*{Allowed Resources}
\begin{itemize}
    \item A text editor for creating files
    \item A browser to check XML format
    \item JSONLint.com to validate JSON files
\end{itemize}

\section*{Restrictions}
\begin{itemize}
    \item No outside assistance is allowed.
    \item Only use skills and knowledge learned in class.
    \item All work must be done independently; violations will result in a 0 grade and be addressed according to the Student Code of Rights and Responsibilities.
\end{itemize}

\end{document}