\documentclass{article}
\usepackage{fancyhdr}

% Header
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{MAD 103 - Data Fundamentals}
\fancyhead[R]{Instructor: Sodiq Shofoluwe \thepage}

\begin{document}

\title{Test \#2 - Important Information}
\author{}
\date{November 7th, 2024}
\maketitle

\section*{Instructions}
Please read the following instructions carefully:

\begin{itemize}
    \item \textbf{Test Date}: Thursday, November 7, 2024, during the regular class time.
    \item \textbf{Content Covered}: Materials from week 3 - 8 (excluding Aggregate Functions).
\end{itemize}

\section*{Test Breakdown}
\begin{itemize}
    \item \textbf{Total Questions}: 14
    \item \textbf{Question Types and Marks}:
    \begin{itemize}
        \item 1 short answer question worth 13 marks
        \item 5 short answer questions worth 3 marks each
        \item 8 short answer questions worth 2 marks each
    \end{itemize}
    \item \textbf{Total Marks}: 44
\end{itemize}

\section*{Guidelines}
\begin{itemize}
    \item Pay attention to the time limit and questions.
    \item \textbf{Question Requirements}: Short answer questions will ask you to create appropriate SQL queries to meet specified requirements.
    \item \textbf{Provided Tables}: Two tables are provided for use in specific questions, with table names mentioned in the questions.
    \item Use a text editor to build queries and test them on \texttt{sql.scweb.ca}.
\end{itemize}

\section*{Allowed Resources}
\begin{itemize}
    \item A text editor to create your queries.
    \item A browser to validate queries on \texttt{sql.scweb.ca} (login required).
\end{itemize}

\section*{Restrictions}
\begin{itemize}
    \item No outside assistance is allowed.
    \item Only use skills and knowledge learned in class.
    \item Answers that do not adhere to class practices will receive a mark of 0.
    \item Work must be completed independently; violations will result in a 0 grade and be addressed according to the Student Code of Rights and Responsibilities.
    \item \textbf{Time Limit}: 50 minutes.
\end{itemize}

\end{document}